Management induced changes of soil organic carbon on global croplands

### 2020-12-14 SOCbudgetPaper/SOCbudget.tex
final update of figure and text

\Author[1]{Kristine}{Karstens}
\Author[1]{Benjamin Leon}{Bodirsky}
\Author[1]{Jan Philipp}{Dietrich}
\Author[2]{Marta}{Dondini}
\Author[1]{Jens}{Heinke}
\Author[2]{Matthias}{Kuhnert}
\Author[1]{Christoph}{Müller}
\Author[1]{Susanne}{Rolinski}
\Author[2]{Pete}{Smith}
\Author[1]{Isabelle}{Weindl}
\Author[1]{Hermann}{Lotze-Campen}
\Author[1]{Alexander}{Popp}

@@

\begin{abstract}
Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth. It is three times larger than the vegetation carbon pool and more than twice as big as the atmospheric carbon pool, when looking into the first meter of the Earth's soil profile. SOC has been depleted by anthropogenic land cover change and agricultural management. However, the latter has so far not been well represented in global carbon stock assessments. While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale. Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils. We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975. Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity. We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
\end{abstract}

@@

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons ({[}portmann\_mirca2000global\_2010{]}) multiple cropping systems (\citep{waha_multiple_2020}), irrigation systems (\citep{jagermeyr_water_2015})) or as stylized scenarios (e.g.~\citep{pugh_simulated_2015}, \citep{lutz_simulating_2019}).

@@

The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historical agricultural management. To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We simulate SOC stocks, dynamics and the SOC debt for 1975--2010. Using a scenario analysis, we decompose the contribution of different management activities.
\newpage

@@

  \begin{table*}[h]
  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
  \tophline
  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
  \middlehline
  \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                      \cite{faostat_faostat_2016}, \\
                                                                      \cite{schaphoff_lpjml4_2018},\\
                                                                      \cite{weindl_livestock_2017} \end{minipage}} & 
                                                      \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                      LG:C generic values according to Table 5.5B,\\
                                                                      5.5C in \cite{calvo_buendia_ipcc_2019},\\
                                                                      crop-specific N:C from \cite{bodirsky_n2o_2012}
                                                                      \end{minipage}} \\
                            & below-ground residues, &  &  \\
                            & manure                &  &  \\ 
                            \hline
  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
                            & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
                                \cite{calvo_buendia_ipcc_2019} (CENTURY parameterization) \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \label{tab:datasourceinputs}
 \belowtable{}
 \end{table*}

@@

Crop production patterns are compiled crop group specific using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see Sect. \ref{sec:landuse}). We calibrate cellular yields with a country-level calibration factor for each crop group to meet historical FAOSTAT production \citep{faostat_faostat_2016}. By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.

@@

Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling. Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see \citep{calvo_buendia_ipcc_2019}).
Depending on the feed system we assume manure to be handled in four different ways:
All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing.
To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields. For manure stored in different animal waste management systems we compiled carbon loss rates depending on the different systems and the associated nitrogen loss rates as specified in \citep{bodirsky_n2o_2012} (see Sect. \ref{append:TableclossAWMS} for more details). We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.

@@

@@

\begin{figure}[h]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland shows high carbon stocks in high yielding areas. (b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state showing different hot spots of SOC dynamics. Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas. (d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the last decades. }\label{fig:SOCmaps}
\end{figure}

In Fig. \ref{fig:SOCmaps}(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010. Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands, to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
Our spatially explicit results show hot spots of SOC losses and gains compared to SOC under potential natural vegetation in two different ways:
1. Absolute SOC changes \(\Delta SOC\) (see Fig. \ref{fig:SOCmaps}(b)) indicate areas with high importance for the global SOC losses. The might be driven by large relative changes (e.g.~in Central Africa) or a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-east Asia). In contrast, large net gains of SOC happen primarily in developed countries of the North according to our results.
2. Relative SOC changes measured as stock changes factors \(F^{SCF}\) (see Fig. \ref{fig:SOCmaps}(c)) are a helpful tool, to measure the impact of human cropping activities. They indicate areas with large difference in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices. Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation. Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have improved, but presumably irrigated croplands at the border to dry, unsuitable areas worldwide as well.

Focusing of the temporal dimension of SOC stock changes (see Fig. \ref{fig:SOCmaps}(c)), global \(\Delta SOC\) has decreased around 15\% over the last couple of decades to \(22\unit{GtC}\). This leads to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\) and considering our estimate of the global SOC stock of \(675\unit{GtC}\) in 2010 to \(0.2\) per mille per year.

\hypertarget{agricultural-management-effects-on-delta-soc}{%
\subsection{\texorpdfstring{Agricultural management effects on \(\Delta SOC\)}{Agricultural management effects on \textbackslash Delta SOC}}\label{agricultural-management-effects-on-delta-soc}}

Figure \ref{fig:SOCscen}(a) reveals the relative impact of management effects by keeping carbon inflows from crop residue returning or manure application constant at the level of 1975 and neglecting the adoption of no-tillage practices. Without changes in management regimes, especially in C inflows from crop residues to the soil, global \(\Delta SOC\) on cropland would be increase at a rate of \(0.1\unit{GtC yr^{-1}}\). The constResidue scenario dominates the effect and is only single-off scenario that still losses SOC compared to natural vegetation at a rate of \(0.06\unit{GtC yr^{-1}}\). Whereas both the constManure and constTillage scenarios, show only small deviations from the historical values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both), the effect of no-tillage gets stronger till 2010.
The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years is visible in Fig. \ref{fig:SOCscen}(b) as well.

\begin{figure}
\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different scenarios deviating from historical argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage). ConstManagement combines all three modifications. Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state. Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
\end{figure}

The results of the sensitivity analysis concerning the initialization of SOC stocks at the beginning of the spin-up phase are shown in Fig. \ref{fig:SOCinit}. The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901. The \(\Delta SOC\) estimation almost halves (from \(~26\unit{GtC}\) to \(~14\unit{GtC}\)) as well as the decrease in of \(\Delta SOC\) for over the last 35 years (from \(~4\unit{GtC}\) to \(~2.5\unit{GtC}\)) for the Inital-natveg scenario.

\begin{figure}
\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901.}\label{fig:SOCinit}
\end{figure}

@@

Figure \ref{fig:FlowFig} illustrates the carbon flows within the agricultural production system on cropland for the year 2010.

\begin{figure}
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows (at arrows) and SOC stocks (in boxes) within the agricultural system for the year 2010 (in MtC): Most important carbon sources on cropland are crop residues.}\label{fig:FlowFig}
\end{figure}

@@

This study shows that spatially explicit and time-variant historical agricultural management data changes estimations of states and trends of global SOC and SOC debt (\(\Delta SOC\)) considerably compared to the often used constant management assumptions. This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks. While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.

@@

The evaluation of SOC stock gains and losses, is complex and has several dimensions as climatic and anthropogenic effects overlap. If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink since 1975 according to our study. However, annual C sequesteration rates of 0.2 per mille are well below the promoted 4 per mille (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.

@@

Furthermore, \citep{sanderman_soil_2017} modeled historical trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. \citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historical data like yield-levels or no-tillage areas were taken into account. Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a substantial decline in SOC debt (less than 1\% change). Historic productivity increases were, however, notably larger. Despite large spatial heterogeneity aggregate yield increase rates are often estimate well above 50\% (\citep{pellegrini_crop_2018}, \citep{ray_recent_2012}, \citep{rudel_agricultural_2009}).

@@

@@

\% latex table generated in R 3.6.3 by xtable 1.8-4 package
\% Mon Dec 14 06:14:28 2020

@@

With regard to the time trend, our study shows the substantial impact of changing management factors on the development of \(F_{SCF}\) as also indicated by the time trend of the SOC debt.

@@

\begin{figure}[h]
\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area. Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time. This makes it hard to pinpoint a specific year for these SOC estimations. In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCglo}
\end{figure}

The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates. SoilGrids (\citep{hengl_soilgrids250m_2017}) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar. Looking at regional results in Fig. \ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas. To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland. Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of \(\Delta SOC\) and \(F^\mathrm{SCF}\).

@@

Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management. It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historical agricultural management on global SOC stocks.

With the extraction of C at harvest and little efforts to increase cropland SOC, these stocks are depleted in comparison to natural vegetation SOC stocks in most regions. The reduction of further deforestation for cropland expansion thus need to be stopped quickly.

However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975. Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils. The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO\textsubscript{2} source. Under the assumption that cropland management is static in time, as typically assumed in other studies on cropland SOC stocks, we also find that cropland soils are a source of CO\textsubscript{2}, building trust in our method.

The estimated increase in cropland SOC is caused by changing agricultural management. The largest contribution to these dynamics comes from increased carbon inputs to soils from crop residues. Crop residue management can still be improved from the SOC perspective, e.g.~by reducing shares of residue burning or further improvements in agricultural productivity. Similarly, other carbon-accumulating practices, such as the application of cover crops need to be included in studies of cropland SOC dynamics, if data on these practices become available.

>>>>

\Author[true]{Kristine}{Karstens}
\Author[1]{Benjamin Leon}{Bodirsky}
\Author[1]{Jan Philipp}{Dietrich}
\Author[2]{Marta}{Dondini}
\Author[1]{Jens}{Heinke}
\Author[2]{Matthias}{Kuhnert}
\Author[1]{Christoph}{Müller}
\Author[1]{Susanne}{Rolinski}
\Author[2]{Pete}{Smith}
\Author[1]{Isabelle}{Weindl}
\Author[1]{Hermann}{Lotze-Campen}
\Author[1]{Alexander}{Popp}

@@

\begin{abstract}
Soil organic carbon (SOC) is one of the largest terrestrial carbon stocks on Earth. The first meter of the Earths soils profile stores three times as much carbon as the vegetation and twice the amount of C in the atmosphere. SOC has been depleted by anthropogenic land cover change and agricultural management. However, the latter has so far not been well represented in global carbon stock assessments. While SOC models often simulate detailed biochemical processes that lead to the accumulation and decay of SOC, the management decisions driving these biophysical processes are still little investigated at the global scale. Here we develop a spatial explicit data set for agricultural management on cropland, considering crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We combine it with the IPCC Tier 2 steady-state soil model to create a half-degree resolution data set of SOC stocks and SOC stock changes for the first 30 cm of mineral soils. We estimate that due to arable farming, soils have lost around 26 GtC relative to a counterfactual natural state in 1975. Yet, within the period 1975--2010 this SOC debt has been decreasing again by a net quantity of 4 Gt SOC, which can be mainly traced back to an increased input of C in crop residues due to higher crop productivity. We also find that SOC is very sensitive to management decisions such as residue returning indicating the necessity to incorporate better management data in soil model simulations.
\end{abstract}

@@

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g.~growing seasons (\citep{portmann_mirca2000global_2010}) multiple cropping systems (\citep{waha_multiple_2020}), irrigation systems (\citep{jagermeyr_water_2015})) or as stylized scenarios (e.g.~\citep{pugh_simulated_2015}, \citep{lutz_simulating_2019}).

@@

The objective of our study is to provide the first global, spatially explicit SOC and SOC debt map that considers spatially explicit and time-variant historic agricultural management. To achieve this objective we create and provide a comprehensive data set of the global gridded management time series data, including crop production levels, residue returning rates, manure application, and the adoption of irrigation and tillage practices. We simulate SOC stocks, dynamics and the SOC debt for 1975--2010. Using a scenario analysis, we decompose the contribution of different management activities.
\newpage

@@

  \begin{table*}[h]
  \caption{Type and data sources for carbon inputs and parameterization to different land-use types }
  \begin{tabular}{p{0.15\textwidth} p{0.20\textwidth} p{0.20\textwidth} p{0.35\textwidth}}
  \tophline
  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
  \middlehline
  \multirow{3}{*}{Cropland} & above-ground residues, & \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                      \cite{faostat_faostat_2016}, \\
                                                                      LPJmL4 \cite{schaphoff_lpjml4_2018},\\
                                                                      \cite{weindl_livestock_2017} \end{minipage}} & 
                                                      \multirow{3}{*}{\begin{minipage}[t]{0.35\columnwidth}\raggedright
                                                                      LG:C generic values according to Table 5.5B,\\
                                                                      5.5C in IPCC \cite{calvo_buendia_ipcc_2019},\\
                                                                      crop-specific N:C from \cite{bodirsky_n2o_2012}
                                                                      \end{minipage}} \\
                            & below-ground residues, &  &  \\
                            & manure                &  &  \\ 
                            \hline
  Natural vegetation        & annual litterfall &  \cite{schaphoff_lpjml4_2018}
                            & \begin{minipage}[t]{0.35\columnwidth}\raggedright\strut 
                                IPCC \cite{calvo_buendia_ipcc_2019} and\\ 
                                CENTURY (\cite{century_model_2000}) \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \label{tab:datasourceinputs}
 \belowtable{}
 \end{table*}

@@

Crop production patterns are compiled crop group specific using half-degree yield data from LPJmL \citep{schaphoff_lpjml4_2018} as well as half-degree cropland patterns (see Sect. \ref{sec:landuse}). We calibrate cellular yields with a country-level calibration factor for each crop group to meet historic FAOSTAT production \citep{faostat_faostat_2016}. By using physical cropland areas in combination with harvested areas, we account for multiple cropping systems as well as for fallow land.

@@

Manure production and management is based on a revised methodology of \citep{bodirsky_n2o_2012} and is presented here due to its central role in soil carbon modeling. Based on the gridded livestock distribution we calculate spatially explicit excretion by estimating the nitrogen balance of the livestock systems on the basis of comprehensive livestock feed baskets \citep{weindl_livestock_2017}, assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios, which range from 10 for poultry up to 19 for beef cattle (for full detail see \citep{calvo_buendia_ipcc_2019}).
Depending on the feed system we assume manure to be handled in four different ways:
All manure originated from pasture feed intake is excreted directly on pastures and rangelands (pasture grazing), deducting manure collected as fuel.
Whereas for low-income countries, we adopt a share of 25\% of crop residues in feed intake directly consumed and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries; middle-income countries see linearly interpolated shares depending on their per-capita income.
For all other feed items, we assume the manure to be stored in animal waste management systems associated with livestock housing.
To estimate the carbon actually returned to the soil, we account for carbon losses during storage, where return shares depend on different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not returned to the fields. For manure stored in different animal waste management systems we compiled carbon loss rates (see calcClossConfinement.R in \citep{bodirsky_mrcommons_2020} for more details) depending on the different systems and the associated nitrogen loss rates as specified in \citep{bodirsky_n2o_2012}. We limit high application shares at \(10\unit{tC\,ha}^{-1}\) to correct for outliers.

@@

Detailed results for the spatially explicit global SOC budget including intermediate results on input data as well as SOC stock results for all scenario runs can be found in \citep{karstens_model_2020}. In the following, the most important results (see \citep{karstens_result_2020} for post-processing script) are summarized.

@@

\begin{figure}[h]
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a): Distribution of total global SOC stocks for the first 30 cm on cropland:  Carbon stocks are large in high yielding areas. (b)+(c): Absolute (b) and relative (c) SOC stocks changes compared to a potential natural state identify different hotspots of SOC dynamics. Whereas absolute losses $\Delta SOC$ are often highest in temperate dry regions, relative losses $F^\mathrm{SCF}$ are often larger in tropical moist areas. (d): $\Delta SOC$ between SOC under historic land use and potential natural vegetation is decreasing over time, meaning net SOC gains on global croplands over the the period 1975--2010. }\label{fig:SOCmaps}
\end{figure}

In Fig. \ref{fig:SOCmaps}(a) we provide the first world map of SOC stocks for the first 30 cm on croplands considering historic management data at the global scale for the year 2010. Values ranging between well over \(100\unit{t ha^{-1}}\) in northern temperate croplands to less than \(5\unit{t ha^{-1}}\) for arid and semiarid croplands.
Our spatially explicit results show hotspots of SOC losses and gains compared to SOC under potential natural vegetation in two complementary ways:
1. Absolute SOC changes \(\Delta SOC\) (see Fig. \ref{fig:SOCmaps}(b)) indicate areas with high importance for the global SOC losses. They might be driven by large relative changes (e.g.~in Central Africa) or by a high natural stock, from which even small relative deviations could lead to substantial absolute losses (e.g.~North-East Asia). In contrast, large net gains of SOC occur primarily in developed countries of the Global North according to our results.
2. Relative SOC changes measured as stock changes factors \(F^{SCF}\) (see Fig. \ref{fig:SOCmaps}(c)) are a helpful metric to analyze the impact of human cropping activities. They indicate areas with large differences in carbon inflows or SOC decay compared to natural vegetation, that may hold potential to be overcome due to improved agricultural practices. Large parts of tropical croplands seem to suffer from low stock changes factors, meaning high relative SOC losses and maybe indicating SOC degradation. Conversely, not only temperate croplands of Central Europe, Japan and western areas of the USA have high stock change factors, but irrigated croplands at the border to dry, unsuitable areas worldwide as well.

The global SOC debt has decreased by about 15\% in the period between 1975 and 2010 to \(22\unit{GtC}\) (Fig. \ref{fig:SOCmaps}(d)). This corresponds to a sequestration rate of \(0.11\unit{GtC yr^{-1}}\). Considering our estimate of the global SOC stock of around \(660\unit{GtC}\) in 1975, global SOC increased by 0.2 per 1000 per year for the period between 1975--2010.

@@

C is sequestered from the atmosphere via plant growth and allocated to three different plant parts (harvest organ, above- and below-ground residues). Whereas harvested organs as well as above ground-residues are taken (partially) from the field to be used for other purposes, below-ground residues (785 MtC in 2010) are directly returned to the field. We split up usage for crop biomass into feed usage and aggregate all other usage types (e.g.~like food, bioenergy and material) into a human demand category. Livestock feed demand for crop harvest and above-ground residues of 1141 MtC is almost equal to the human demand of 1144 MtC. Whereas large parts of feed intake are recycled to the soils via manure (C input from manure at 384 MtC), we assume the carbon demanded from humans (ending up as e.g.~compost, night soil and sewage) is not recycled to soils. Besides manure C and below-ground residues, above-ground residues form the largest C input to the soil with 1200 MtC returned to the field in 2010. However, around 60\% of C decomposes before its integrated into soils at the the litter-soil barrier. Due to the different C composition, proportional more C enters the slow pool from manure compared to crop residue. According to our model results, land-use change dynamics led to a C transfer from cropland to natural vegetation of 58 MtC in 2010. 4764 MtC sequestered by crop plants face 3897 MtC released within the agricultural system. Accounting for SOC transfer, SOC increase under cropland is around 809 MtC.

\begin{figure}[h]
\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/CarbonBudget} \caption{Global carbon flows within the agricultural system for the year 2010 (in MtC): Carbon is first photosynthesise by crop plants and than used depending on the plant part by humans for feed of livestock and various other usages subsumed under human demand. After accounting for losses within the agricultural system three major C inputs are applied to croplands: manure, above- and below-ground residues. Large parts of C, however, get mineralized on field before entering the soil. Additionally, C is transferred to and from agricultural soils via land-use change to and from natural vegetation. Finally, SOC gets mineralized and flows back into the atmosphere.}\label{fig:FlowFig}
\end{figure}

\hypertarget{agricultural-management-effects-on-soc-debt}{%
\subsection{Agricultural management effects on SOC debt}\label{agricultural-management-effects-on-soc-debt}}

We analyze the relative impact of individual management aspects by comparing the actual historic management scenario with counterfactual scenarios where individual management aspects are kept static at the 1975 values (Figure \ref{fig:SOCscen}(a)). Without changes in management regimes, the global \(\Delta SOC\) on cropland would be increasing at a rate of \(0.1\unit{GtC yr^{-1}}\). As shown by the constResidue scenario, changes in residue return rates dominate the management effects. Without the historic increase in residue returning, the global \(\Delta SOC\) would still increase at a rate of \(0.06\unit{GtC yr^{-1}}\). Both the constManure and constTillage scenarios show only small deviations from the historic values (sequestration rate of \(0.09\unit{GtC yr^{-1}}\) for both). The effect of no-tillage has been particuarly strong since the 1990s.
The strong impact of almost doubling C inputs from crop residue biomass over a period of 35 years on agricultural SOC stocks is shown in Fig. \ref{fig:SOCscen}(b).

\begin{figure}
\includegraphics[width=14cm]{../ResultNotebooks/Output/Images/scenario_horiz} \caption{(a) Global $\Delta SOC$ in GtC for different management scenarios: The stylized scenarios devate from historic argicultural management by holding effects of carbon inflows from residues (constResidues), manure (constManure) constant or neglecting adoption of no-tillage practices (constTillage). ConstManagement combines all three modifications. Note that $\Delta SOC$ is defined as the difference of SOC under land-use compared to a natural vegetation state. Figure (b) shows the carbon inflows from crop residue and manure, underlining the strong impact of residues for SOC stock and SOC stock changes.}\label{fig:SOCscen}
\end{figure}

Our sensitivity analysis shows that the management impact is robust to the initialization of SOC stocks at the beginning of the spin-up phase (Fig. \ref{fig:SOCinit}). The Initial-natveg scenario initializes the start year with the steady-state SOC under potential natural vegetation for all land-use types, compared to the default assumption (Initial-lu), which assumes landuse-type specific steady-state SOC stocks in 1901. The \(\Delta SOC\) estimation almost halves (from \textasciitilde26 GtC to \textasciitilde14 GtC), as does the decrease in \(\Delta SOC\) for the period of 1975--2010 (from \textasciitilde4 GtC to \textasciitilde2.5 GtC) for the Inital-natveg scenario.

\begin{figure}
\includegraphics[width=10cm]{../ResultNotebooks/Output/Images/scenario_init} \caption{Global $\Delta SOC$ for different SOC initialization choices in the start year 1901: Starting the spin-up phase with natural steady-state SOC under vegetation without any human cropping activities (Initial-natveg) lead to a smaller $\Delta SOC$ in 1975 and a less steeper increase till 2010, as compared to initialzing with steady-state SOC stocks under historic land-use.}\label{fig:SOCinit}
\end{figure}

Global SOC stocks range between 645--700 GtC for different parameterization of the natural litterfall, whereas \(\Delta SOC\) vary 19-24 GtC for the year 2010 (see Fig. \ref{fig:SOClitter}. The increase of \(\Delta SOC\) during the period of 1975--2010 is independent of the parameterization choice at \textasciitilde4 GtC.

@@

This study shows that spatially explicit and time-variant historic agricultural management data considerably alter estimates of the states and trends of SOC compared to the often used constant management assumptions. This result remains robust to variations of central model parameters and variations in the initialization of SOC stocks. While land cover change has depleted SOC stocks and increased the SOC debt, our analysis points out that the high increases in agricultural productivity may have even led to a net reduction of the SOC debt since 1975.

@@

The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap. If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see \citep{pugh_simulated_2015}), global croplands have acted as a carbon sink since 1975 according to our study. However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 (\citep{minasny_soil_2017}), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.

@@

Furthermore, \citep{sanderman_soil_2017} modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. \citep{pugh_simulated_2015} considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account. Their study moreover concludes that crop productivity-gains (increasing yield levels by 18\% in their simulations) do not lead to a~substantial decline in SOC debt (less than 1\% change). Historic productivity increases were, however, notably larger. Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50\% (\citep{pellegrini_crop_2018}, \citep{ray_recent_2012}, \citep{rudel_agricultural_2009}).

@@

Looking on historic cropland intensification rates of above 50\% (\citep{rudel_agricultural_2009}), a large fraction of increased C input from residue biomass is attributed to productivity improvements.

@@

\% latex table generated in R 3.6.3 by xtable 1.8-4 package
\% Mon Dec 14 23:33:09 2020

@@

With regard to the time trend, our study shows the substantial impact of changing management factors on the development of stock change factors as also indicated by the time trend of the SOC debt.

@@

\begin{figure}[h]
\includegraphics[width=8cm]{../ResultNotebooks/Output/Images/glo_comparisonfigure} \caption{Modeled as well as data based estimation for global SOC stock in GtC for the first 30 cm of soil aggregated over all land area: Note that SoilGrids, GSOC and WISE do not consider land-use as well as changes over time and rely on soil profile data gather over a long period of time. This makes it hard to pinpoint a specific year for these SOC estimations. In this context they will be compared to modeled data from LPJmL4 for potential natural vegetation and this study for the year 2010.}\label{fig:SOCglo}
\end{figure}

The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates. SoilGrids (\citep{hengl_soilgrids250m_2017}) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar. Looking at regional results in Fig. \ref{fig:SOCreg}, our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas. To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland. Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.

@@

Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management. It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.

Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models. That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data. Our data-set and the MADRaT package (\citep{dietrich_madrat_2020}) constitute a starting point for building comprehensive data sets on agricultural management aspects.

Our study again highlights that the expansion of croplands is still a major source of CO\textsubscript{2} emissions -- not only by the removal of vegetation, but also by a slow depletion of soils. Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.

However, our results also indicate that the changes in cropland management have led to increased SOC stocks in global cropland soils as a continuous trend since 1975. Even more, this trend of improved management on existing croplands more than overcompensates the depletion of SOC stocks on newly converted soils. The finding of this recovery of cropland SOC stocks challenges the assumption that cropland soils are a CO\textsubscript{2} source. Only under the assumption that cropland management is static over time, as typically assumed in other studies on cropland SOC stocks, we can reproduce their finding that cropland soils are a source of CO\textsubscript{2}. The estimated increase in cropland SOC is therefore caused by changing agricultural management, with the largest contribution stemming from increased carbon inputs to soils from crop residues.

Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks. Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand (\citep{popp_land-use_2014-1}) -- and thereby usually further reduce the SOC debt on croplands.

While further crop productivity gains lead to co-benefits for SOC stocks, they are likely not enough to help meeting ambitious climate mitigation targets. Our findings on sequestration rates are still more than an order of magnitude lower than promoted by the 4 per 1000 initiative.
